## Model yeast cell replication from timelapse image

Using the image [YDA306_AGA1y_PRM1r_Mating](https://idr.openmicroscopy.org/webclient/?show=image-3491334) from IDR to model the replication of yeast cells.

### Load libraries

```{r}
library(romero.gateway)
library(EBImage)
```

### Connect to server

```{r}
server <- OMEROServer(host = 'wss://idr.openmicroscopy.org/omero-ws', port = 443L, username='public', password='public')
server <- connect(server)
paste('Successfully logged in as', server@user$getUserName())
```

### Load image

```{r}
imageId <- 3491334
image <- loadObject(server, "ImageData", imageId)
```

### Set \# of timepoints and interval

(this information could be retrieved from the server too, but lets keep it simple)

```{r}
tmax <- 42  # number of timepoints
tInMinutes <- 5  # separation between timepoints in minutes
```

### Get the pixel values and display image

```{r}
t <- as.integer(tmax / 2) # use half-time point
pixels <- getPixelValues(image, 1, t, 2)
ebi <- EBImage::Image(data = pixels, colormode = 'Grayscale')
img <- normalize(ebi)
EBImage::display(img)
```

### Threshold the image

```{r}
threshImg <- thresh(img, w=20, h=20, offset=0.05)
threshImg <- medianFilter(threshImg, 3)
threshImg <- fillHull(threshImg)
threshImg <- bwlabel(threshImg)
EBImage::display(colorLabels(threshImg))
```

### Count the number of cells

```{r}
count <- range(threshImg)
count[2]
```

### Wrap up thresholding and counting into a function

```{r}
countCells <- function (t) {
  pixels <- getPixelValues(image, 1, t, 2)
  ebi <- EBImage::Image(data = pixels, colormode = 'Grayscale')
  img <- normalize(ebi)
  threshImg <- thresh(img, w=20, h=20, offset=0.05)
  threshImg <- medianFilter(threshImg, 3)
  threshImg <- fillHull(threshImg)
  threshImg <- bwlabel(threshImg)
  count <- range(threshImg)
  as.integer(count[2])
}
```

### Analyse all time points

```{r}
timepoints <- (1:tmax) # analyze every time point
#timepoints <- seq(1,tmax,7) # only use every seventh time point (faster)
df <- data.frame(TimePoint=timepoints)
df$CellCount <- sapply(df$TimePoint, countCells)
df$TimePoint <- sapply(df$TimePoint, function(x) x * tInMinutes)
df
```

### Plot the result

```{r}
plot(df$TimePoint, df$CellCount, ylab = "Number of cells", xlab = "Time (in min)")
```

### Fit a model

```{r}
model <- lm(log(df$CellCount)~ df$TimePoint)
print(summary(model))
```

### Lets see how well it fits visually

```{r}
plotTime <- as.list(df[,1])
predCounts <- exp(predict(model,plotTime))
plot(df$TimePoint, df$CellCount, ylab = "Number of cells", xlab = "Time (in min)")
lines(plotTime, predCounts,lwd=2, col = "red")
```

### Disconnect from the server again

```{r}
disconnect(server)
```
